01-Travel-time-calculation

### 2022-04-08 00-OD-preparation.Rmd
Calculated travel times for car, walk, bike, bus for all OD combinations in Madrid. Completed some preliminary comparisons to survey travel times.

zn <- st_read("data-raw/zn.shp")
```

## Filtering OD trips, home to work

@@

Well, the majority of trip frequency is "A diario, de lunes a viernes" (Daily, from Monday to Friday - 1)
```{r}
hist(OD_v_Oc_Dt$VFRECUENCIA)
```
The average destination arrival time (Destino. Hora llegada) is:
```{r}

timeD <- as.POSIXct(OD_v_Oc_Dt$VDESHORAFIN, format="%H%M") #formating the column
timeD <- format(timeD, "%H:%M") # removing date so we just how time

seconds_to_period(mean(period_to_seconds(hm(timeD)), na.rm=T)) # finding mean

```
9:48am arrival.. not very helpful.

Let's see the distribution:
```{r}
hist(OD_v_Oc_Dt$VDESHORAFIN %>% as.numeric(), breaks = 24)
```
So we see that 8AM, and 9AM arrival times are most frequent.

@@

Let's see how long trips are (by distance):
```{r}
t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE <= 60) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t)
```

@@

## Saving data
```{r}
save(zn, file = "data/zn.RData")
```

>>>>

zn <- st_read("data-raw/zn.shp")
```
Importing the spatial object which corresponds to the low-emission-zones
```{r}
#XXX
```

## Organizing the trip level data

```{r}
#adding travel time... we don't even need r5r! we just need to decide how to remove A polluting cars
OD_v <- OD_v %>% mutate(
  TIEMPO_VIAJES = 
                  as.numeric((as.POSIXct(OD_v$VDESHORAFIN, format="%H%M") -
                  as.POSIXct(OD_v$VORIHORAINI, format="%H%M"))/60))

```

## Exploring the data - filtering OD trips, home to work

@@

Well, the majority of trip frequency is "A diario, de lunes a viernes" (Daily, from Monday to Friday - 1)
```{r}
hist(OD_v_Oc_Dt$VFRECUENCIA)
```
The average destination arrival time (Destino. Hora llegada) is:
```{r}
Dtrips <- as.POSIXct(OD_v$VDESHORAFIN, format="%H%M") %>% format("%H:%M") # formatting and removing date so we just how time
seconds_to_period(mean(period_to_seconds(hm(Dtrips)), na.rm=T)) # finding mean
```
9:48am arrival.. not very helpful.

Let's see the distribution:
```{r}
hist(OD_v_Oc_Dt$VDESHORAFIN%>% as.numeric(), breaks = 24)
```
So we see that 8AM, and 9AM arrival times are most frequent.

@@

Let's see how long are short-trips are (by distance, assuming in km units):
```{r}
t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE <= 60) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t)
```

As you can see though there are some long distance trip (greater than 60km). There are 101 trips which are between 60km to 600km and 1 trip greater than 600km. For perspective though, these trips make up onl 0.3% of total trips taken ( ~100 trips out of ~28,000 trips).
```{r}
t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE >= 60 & DISTANCIA_VIAJE <= 600) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t, breaks = 50)

t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE >= 600) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t, breaks = 50)
```

Let's check the primary mode: 
```{r}
hist(OD_v_Oc_Dt$MODO_PRIORITARIO)
```

@@

## Saving data
```{r}
save(zn, file = "data/zn.RData")
save(OD_v, file = "data/OD_v.RData")
```

### 2022-04-08 01-Travel-time-calculation.Rmd
Calculated travel times for car, walk, bike, bus for all OD combinations in Madrid. Completed some preliminary comparisons to survey travel times.

# --- Formating the data
```{r}

Format the origin and destination points (ID, Long, Lat). In this case, all origins and all destinations are the same (i.e., 1259 zones so 1259 origins and 1259 destinations, so 1,585,081 OD travel times from each origin to destination):
```{r}
# # I was having an error running st_centroid, this line fixes it.
# sf::sf_use_s2(FALSE)

@@

```{r}
#download and import GTFS (Madrid transit data)
# download.file(url = "https://transitfeeds.com/p/emt-madrid/212/latest/download", destfile = file.path(r5_path, "MAD_transit.zip"), mode = "wb")

@@

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_MAD,
                          origins = Orig_chunk,
                          destinations = OD_all,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2022-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "UTC+2")),
                          max_trip_duration = 120)

@@

>>>>

# --- Formating the data
Format the origin and destination points (ID, Long, Lat). In this case, all origins and all destinations are the same (i.e., 1259 zones so 1259 origins and 1259 destinations, so 1,585,081 OD travel times from each origin to destination):
```{r}
# # I was having an error running st_centroid, this line fixes it.
# sf::sf_use_s2(FALSE)

@@

```{r}
#download and import GTFS (Madrid transit data)
download.file(url = "https://transitfeeds.com/p/emt-madrid/212/latest/download", destfile = file.path(r5_path, "MAD_transit.zip"), mode = "wb")

@@

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_MAD,
                          origins = Orig_chunk,
                          destinations = OD_all,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2022-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "CEST")),
                          max_trip_duration = 120)

@@

## Transit travel time
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(OD_all)/chunksize)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_MAD,
                          origins = Orig_chunk,
                          destinations = OD_all,
                          mode = c("TRANSIT", "WALK"),
                          departure_datetime = as.POSIXct(strptime("2022-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "CEST")),
                          max_trip_duration = 120)

output_OD_transit <- as.data.frame(output_df)

save("output_OD_transit", file = "data-raw/intermed/ttresults/output_OD_transit.Rdata")
#I only made the cut off 60 minutes, so if some parcels do not have a travel time, rerun this for a 120 min cut off. The following are 120mins cutoff
```

## Walk travel time
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(OD_all)/chunksize)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_MAD,
                          origins = Orig_chunk,
                          destinations = OD_all,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2022-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "CEST")),
                          max_trip_duration = 60)

output_OD_walk <- as.data.frame(output_df)

save("output_OD_walk", file = "data-raw/intermed/ttresults/output_OD_walk.Rdata")
```

## Bike travel time
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(OD_all)/chunksize)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_MAD,
                          origins = Orig_chunk,
                          destinations = OD_all,
                          mode = c("BICYCLE"),
                          departure_datetime = as.POSIXct(strptime("2022-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "CEST")),
                          max_trip_duration = 60)

output_OD_bike <- as.data.frame(output_df)

save("output_OD_bike", file = "data-raw/intermed/ttresults/output_OD_bike.Rdata")
```
